A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-07-27 text/manuscript.Rmd
Pull survival text out into separate survival.Rmd file, add a Makefile to build the document from the command line

If our molecular pathway disruption classes represent distinct biological states
apart from tissue of origin, they may have different prognoses within cancer
types or across all cancers. Kaplan-Meier plots reveal large differences between
cancer types as expected, and some modest differences between our classes (not
shown). These analyses are profoundly limited by confounding factors of age and
stage at diagnosis, sex, ethnicity, and tissue-specific disease progression, to
name the obvious ones. To explore these ideas, we tested models of survival
using public longevity data from the CDC, accounting for age, gender and
ethnicity. To build on this relatively accurate model of baseline risk (see
methods), we assume that individual cancer type captures some variation not
accounted for in the longevity data ( _e.g._ all tissue-specific factors,
including stage [see methods]), thus allowing us to avoid introducing
uncertainty by modeling each of those factors explicitly. Therefore, our model
estimates the effect of cancer type and class-specific cancer effects
independently. We found that cancer types, as expected, have a range of
prognoses relative to the general population, as shown in **figure 5a**. The
least deadly was prostate cancer (PRAD), and the most deadly was glioblastoma
(GBM) which has a risk multiplier of between 15 and 22 relative to the
background risk of death in the population. Among the deadliest cancers outside
of GBM were stomach (STAD), melanoma (SKCM) and pancreatic (PAAD) cancers. Apart
from these top 4 cancers which ranged from 6-22 in relative risk of age-adjusted
death, the remaining cancer types ranged from about 1 to 5 in magnitude
(**figure 5a**).

![](../figures/fig6B.png)
![](../figures/fig6C.png)
<p style="FONT-SIZE:10px; COLOR:#000000; LINE-HEIGHT:12px;">**Figure 5: relative survival of pathway clusters. a) Posterior distributions of estimated relative age-dependent survival by cancer type. b) Posterior distributions of estimated class-dependent survival for each cancer type. See methods for model details.</p>

Our estimates of tissue-specific class effects, in contrast, ranged from 0-3,
reflecting that some classes are either less deadly or more deadly than other
classes within each cancer types, where the null hypothesis corresponds to an
effect size=1 (**figure 5b**). For several cancers ( _e.g._ PRAD, kidney
chromophobe (KICH), diffuse large B-cell lymphoma (DBLC), thyroid cancer (THCA))
the posterior estimates are largely indestinguishable from the prior, reflecting
that either there were too few mortalities in the data to make an estimate (as
expected for PRAD and THCA) or two few samples period. We did not observe
class-specific trends that held true across cancer types, which could result
from different cancers having different standards of care for example. In
support of this interpretation, we also tested a factored model which considered
both cancer type and each class independent of cancer type (**supplementary
figure S7??**). Though we were able to successfully fit these models, they
exhibited much greater uncertainty, suggesting an inferior specification.
(<mark>Dan are there additional arguments to be made for the inferiority of the
factored model, here?</mark>)

```{r survival-model, eval=FALSE}

```

![**Figure 6B**](../figures/fig6B.png)

![**Figure 6C**](../figures/fig6C.png)

@@

All code for producing the analyses and figures herein are included in this
fully reproducible manuscript. In addition, the R markdown files and all other
models are available from repositories in the following locations on the
distributed version software hub GitHub.

>>>>

```{r child='survival.Rmd'}
```

@@

All code for producing the analyses and figures herein are included in this
fully reproducible manuscript. In addition, the R markdown files and all other
models are available from
[our repositories](https://github.com/dennishazelett/TTmanu) in the following
locations on the distributed version control site GitHub
